---
title: "Homework 5"
author: "David"
date: "2024-11-09"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

```

## Problem 1

Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years and that birthdays are uniformly distributed over the year.
Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.

```{r}
bdays_sim = function(n) {

  bdays = sample(1:365, size = n, replace = TRUE)

  duplicate = length(unique(bdays)) < n
  
  return(duplicate)

}

```


Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. Make a plot showing the probability as a function of group size, and comment on your results.

```{r}
sim_res =
  expand_grid(
    n = 2:50,
    iter = 1:10000
  ) |>
  mutate(res = map_lgl(n, bdays_sim)) |>
  group_by(n) |>
  summarise(prob = mean(res))

sim_res |>
  ggplot(aes(x = n, y = prob)) +
  geom_line() +
  theme_minimal() +
  labs(
    title =  "Probability that at least two people have same birthday",
    x = "Number of people",
    y = "Probability"
  )

```

Probability of two people having same day of birthday gradually increases as number of people increase. With the speed of increase going up from 2 to around 25 and going down with number of people greater than 25. Probability eventually reaches close to 1.00 .

## Problem 2

```{r}
norm_test = function(mu) {
  
  norm_distribution = rnorm(30, mean = mu, sd = 5)
  
  table =
    t.test(norm_distribution, mu = 0) |>
    broom::tidy()
  
  return(table)
}

norm_sim =
  expand_grid(
    mu = 0:6,
    iter = 1:5000
  ) |>
  mutate(stat_table = map(mu, norm_test)) |>
  unnest(stat_table) |>
  janitor::clean_names() |>
  select(mu:conf_low) 

```

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of mu on the x axis.

```{r}
norm_sim |>
  mutate(
    h = case_when(
      p_value < 0.05 ~ 1,
      p_value >= 0.05 ~ 0
    )
  ) |>
  group_by(mu) |>
  summarise(prob = sum(h) / 5000) |>
  ggplot(aes(x = mu, y = prob)) +
  geom_line() +
  theme_minimal() +
  labs(
    title = "Probability of Rejecting H0 Hypothesis in t Test",
    x = "Mu",
    y = "Probability"
  )

```

As effect size increases, the power of the test also increases. This is because larger effect sizes create more distinct differences or stronger associations, making it easier for statistical tests to detect them. For instance, in a study with a small sample size, a large effect size may still yield high power, while a small effect size may result in low power, risking the failure to detect a true effect.


Make a plot showing average estimate of mu on y axis and true mu on x axis. Add another layer shwoing the average estimate of mu only in samples for which the null hypothesis was rejected.

```{r}
norm_sim |>
  group_by(mu) |>
  summarise(
    mu_mean_all = mean(estimate),
    mu_mean_1 = mean(estimate[p_value < 0.05])
  ) |>
  ggplot(aes(x = mu)) +
  geom_line(aes(y = mu_mean_all, color = "All Samples")) +
  geom_line(aes(y = mu_mean_1, color = "Rejected Samples")) +
  theme_minimal() +
  labs(
    title = "Average Mu in Tests against Real Mu",
    x = "Mu",
    y = "Average Mu in Tests"
  ) +
  scale_color_manual(
    name = "Sample Type",
    values = c("All Samples" = "blue", "Rejected Samples" = "green4"),
    labels = c("All Samples", "Rejected Samples")
  )

```

Across tests where the null hypothesis is rejected, the sample average of mu is often greater than the true value of mu due to selection bias. Only samples with larger estimates of mu are likely to achieve statistical significance, especially for smaller values of mu. This leads to an inflated average estimate in the subset of rejected tests, which does not accurately reflect the true mu.

## Problem 3

```{r}
homicide = read_csv("homicide-data.csv") |>
  janitor::clean_names()

```

The dataset has 52179 observation and 12 variables, including report date, victim information, location of homicide and disposition status.

```{r}
homicide =
  homicide |>
  mutate(city_state = paste(city, state, sep = ",")) |>
  group_by(city_state) |>
  summarise(
    total = n(),
    unsolved = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )
```

Function of prop.test used in below chart and plot. (applied **broom::tidy()** here instead of below code chunks)

```{r}
test = function(unsolved, total){
  
  table =
    prop.test(unsolved, total, p = NULL, alternative = "two.sided", conf.level = 0.95, correct = TRUE) |>
    broom::tidy() |>
    select(estimate, conf.low, conf.high)
  
  return(table)
}
```

For the city of Baltimore, MD, estimate the proportion of homicides that are unsolved; save the output as an R object, pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
homicide |>
  filter(city_state == "Baltimore,MD") |>
  mutate(table = map2(unsolved, total, \(x, y) test(unsolved = x, total = y))) |>
  pull(table) |>
  knitr::kable(
    col.names = c("Estimate", "Lower limit of 95% CI", "Upper limit of 95% CI")
  )
```

Now for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. List columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
homicide =
  homicide |>
  group_by(city_state) |>
  summarise(table = map2(unsolved, total, \(x, y) test(unsolved = x, total = y))) |>
  unnest(table) |>
  mutate(
    estimate = round(estimate, 3),
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3),
    CI = paste(conf.low, conf.high, sep = ",")
  )

homicide |>
  select(-conf.low, -conf.high) |>
  knitr::kable(
    col.names = c("City, State", "Estimate", "95% CI")
  )

```

**Tulsa,AL only has one case so the estimate might not be correct.**

Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r}
homicide |>
  ggplot(aes(x = fct_reorder(city_state, estimate), y = estimate)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "black") +
  coord_flip() +
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimate of Proportion Unsolved"
  ) +
  theme_minimal()

```

**Tulsa,AL only has one case so the estimate might not be correct.**
